### 2018-09-25 manuscript/templates4Authors/latex_template/author.tex
starting manuscript;figures

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% track changes
\usepackage[normalem]{ulem}
\usepackage[dvipsnames]{xcolor}
\newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}

\begin{document}

\title*{Prediction of functional markers of mass cytometry data via
  deep learning}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\thanks{Both authors contributed equally to this manuscript}
\author{Claudia Sol\'is-Lemus, Xin Ma, Maxwell Hostetter II, Suprateek
  Kundu, Peng Qiu, Daniel Pimentel-Alarc\'on}
%\author{Xin Ma, Claudia Sol\'is-Lemus, Maxwell Hostetter II, Suprateek
%  Kundu, Daniel Pimentel-Alarc\'on}
%\author{xxx\thanks{work the same}, xxs$^*$}% \footnotemark[1]}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Claudia Sol\'is-Lemus \at Emory University, Atlanta, GA, \email{csolisl@emory.edu}
\and Xin Ma \at Emory University, Atlanta, GA,
\email{xin.ma@emory.edu}
\and Maxwell Hostetter II \at Georgia State University, Atlanta, GA,
\email{mhostetter1@student.gsu.edu}
\and Suprateek Kundu \at Emory University, Atlanta, GA,
\email{suprateek.kundu@emory.edu}
\and Peng Qiu \at Georgia Institute of Technology, Atlanta, GA,
\email{peng.qiu@bme.gatech.edu}
\and Daniel Pimentel-Alarc\'on \at Georgia State University, Atlanta, GA,
\email{pimentel@gsu.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract{add here}

\section{Introduction}

% from Peng Qui PQ.docx: ---
Multiparametric single-cell analysis has advanced our understanding of
diverse biological and pathological processes, providing insights into
cellular differentiation, intracellular signaling cascades and
clinical immunophenotyping. Modern flow cytometers typically provide
simultaneous single-cell measurements of up to 12 fluorescent
parameters in routine cases, and analysis of up to 30 protein
parameters has been recently made commercially available. In addition,
a next-generation mass cytometry platform (CyTOF) has become
commercially available, which allows routine measurement of 50 or more
protein markers.

Despite the technological advances in acquiring an
increasing number of parameters per single cell, approaches for
analyzing such complex data lag behind. The existing approaches are
often subjective and labor-intensive. For example, the widely used
gating approach identifies cell types by user-defined sequences of
nested 2-D plots. There have been efforts to develop clustering
algorithms (eg, flowMeans, flowSOM, X-shift) and dimension reduction
algorithms (eg, SPADE, tSNE, Scaffold). However, there is still huge
space for developing new methods to ask new questions in this field.
% ---

Recently, deep learning models are revolutionizing the fields of
precision medicine, data mining, astronomy, human-computer
interactions, among many others, by becoming a major discovery force
in science due to the unprecedented accuracy in prediction.  Here, we
explore neural network models to predict functional markers (internal
phosphoproteins) with identify markers (cell surface proteins).

On one side, identify markers define a cell type (e.g., B cell, T
cell, monocytes, MSC), and on the other side, expressions of
functional markers identify the cell's activity (e.g., quiescent,
secreting cytokines, proliferating, apoptosis). Since CyTOF technology
allows for the simulteanous measurement of a large number of protein
markers, most CyTOF studies measure both identity markers and
functional markers, providing data for supervised learning tools, like
neural networks.

\begin{itemize}
\item Mass cytometry (CyTOF) is an emergent technological development
  for high-dimensional multiparameter single cell analysis.
\item CyTOF provides unprecedented multidimensional immune cell
  profiling and has recently been applied to characterizing peripheral
  blood cells, Natural Killer cells in viral infections, skin cells,
  cells in celiac disease, responding phenotypes in cancer and even
  holds the promise of examination of solid tumors (add references)
\item Among the features per cell, there two classes: surface markers
  (which identify the type of cell) and functional markers (which
  identify the function of the cell)
\item Collecting the functional markers is time-consuming and
  computationally expensive (yes?), so it is desirable to devise a
  prediction scheme that can allow us to predict the funcional markers
  from the surface markers
\item Deep learning methods have achieved outstanding performance in
  various computational tasks, such as image analysis, natural
  language processing, and pattern recognition (references)
\item Moreover, recent efforts to use deep learning approaches in
  genomics and biomedical applications show their flexibility for
  handling complex problems (Cires ̧an et al., 2013; Cruz-Roa et al.,
  2014; Denas and Taylor, 2013; Fakoor et al., 2013; Leung et al.,
  2014).
\item in each FCM and CyTOF run we typically collect approximately $10^5$
  to $10^6$ cells, so that the number of instances (cells) is several
  orders of magnitude larger than the number of variables (up to 50
  markers). Therefore, developing deep learning approaches to analyze
  cytometry data is very promising.
\item We present DeepCyTOF, an integrated deep learning domain
  adaptation framework, which employs one manually gated reference
  sample and utilizes it for automated gating of the remaining samples
  in a study.
\item We use this method in the dataset bla
\item We compare the NN approach to other usual prediction methods
  like regression-based and xxx (Suprateek is checking this), and show
  how the NN approach improves the prediction by the incorporation of
  non-linear and interaction effects.
\item We show the improvement driven by NN when the effects are
  nonlinear as a powerful alternative to the usual regression techniques.
\end{itemize}

\section{Data}

In a dataset we previously published
[Bendall et al 2011; Qiu et al 2011], we measured 13 identity
(surface) markers and 18 functional markers simultaneously.  This
measurement was performed on 5 health donor bone marrow samples
untreated, and also on 24 health bone marrow samples treated by 24
different drugs.  For the moment, let’s focus on the first 5 samples.

Given the data from 1 of the 5 samples, we have observations (data
points) for roughly 250,000 cells. This 250,000*(13+18) data can be
used to build a machine learning model to predict the 18 function
markers based on the 13 identity markers. In this dataset, we roughly
know that the 13 identity markers can be used to define ~25 cell
types, and we have manually created cell type label for each cell. One
machine learning question can be predicting the cell type label based
on the 13 surface markers, however, this is not the question we are
asking here. The questions we want to ask are:

1. If we use 1 sample
to train a deep learning model which use identity markers to predict
functional markers, can this model be used (generalized) to predict
functional markers in the other 4 untreated samples?

2. Based on the
trained deep learning model, can we try to identify cell clusters? And
will our cell clusters agree with the well-accepted cell types in the
literature.

3. If we focus cells belonging to the same known cell
type, and examine the distribution of functional markers and the
correlation with the subtle variations of the identity markers among
cells of this type, are we able to see hints that this cell type may
be further divided into subtypes?

The data is highly complex, violating the linearity assumptions of
linear regression, see plot, even after the standard transformation
for CyTOF data (arcsinh with cofactor=5). \ref{exploratory}

\begin{figure}
\includegraphics[scale=0.3]{figures/surface-hist.pdf}
\includegraphics[scale=0.3]{figures/functional-hist.pdf}
\includegraphics[scale=0.3]{figures/scatterplot.pdf}
\caption{Exploratory plots of surface and functional markers. The
  histograms show a biased pattern, and the scatterplot shows non
  linearity, crucial assumptions in regression.}
\label{exploratory}
\end{figure}

\section{Materials and Methods}

We fit a neural network model to predict functional markers from
surface markers, and compare its performance to three classical
statistical methods like linear regression (unpenalized and
penalized), and random trees. Due to computational time constraints,
we could not fit a support vector regression in the dataset. We
compared the performance of the four approaches by computing the mean
square error of the predicted responses.

To fit the models, we divided the data (1,223,228 cells in 18
functional markers (responses) and 15 surface markers (predictors))
into three subsets: 250,000 rows as training data, 250,000 rows as
validation data and 223,228 rows as testing data. The training data
was used to fit each of the four models. The validation data was used
to determine the best setup (tuning parameters) of each model in terms
of mse, and the test data was used to compare the mse across methods.

We used the vector mse (put formula), and the 18 individual mse's (put
formula) on the test data to compare across methods.

\begin{equation}
MSE_{vec} = \frac{1}{n} \sum_{i=1}^n ||\hat{Y}_i - Y_i ||^2_2
\label{vectorMSEformula}
\end{equation}

where $\hat{Y}_i \in \mathbf{R}^{18}$ is the predicted vector of
responses for individual $i$, and $Y_i \in \mathbf{R}^{18}$ is the
observed vector of responses for individual $i$.

The individual MSE was then defined for predictor $k$ as
\begin{equation}
MSE_{(k)} = \frac{1}{n} \sum_{i=1}^n (\hat{Y}_{k,i} - Y_{k,i} )^2
\label{MSEformula}
\end{equation}
where $\hat{Y}_{k,i} \in \mathbf{R}$ is the $k^{th}$ predicted
response ($k=1,\cdots,18$) for individual $i$, and $Y_{k,i} \in
\mathbf{R}$ is the $k^{th}$ observed response for individual $i$.

% https://github.com/pluskid/Mocha.jl/blob/master/docs/user-guide/solver.rst
\vspace{0.25cm}
\noindent \textbf{Neural network model} We tested different network
architectures, activation functions, regularization coefficients,
solver methods, momentum policies, and learning rates. The best
network has four layers (see figure xxx) with 90, 90, 45 and 45
nodes. The network uses hyperbolic tangent as activation function,
regularization coefficient 0.0001, momentum policy at 0.8, learning
rates of (0.05, 0.01, 0.0001, 0.75).

\vspace{0.25cm}
\noindent \textbf{Linear regression (unpenalized/penalized)} We fit standard
linear regression, as well as the penalized version with LASSO penalty
under different penalization parameters. We used the
\textit{ScikitLearn} julia wrapper, with default settings.

\vspace{0.25cm}
\noindent \textbf{Decision tree and random trees} We fit one decision
regression tree with \textit{ScikitLearn} julia wrapper, with default
settings. We compared the performance of the ``mse'' criterion and the
Friedman's improvement score, deciding on the former (``mse'') which
is the default setting. We did not constraint the maximum depth of the
tree, and used as 2 the minimum number of samples required to split an
internal node. We did not constraint the maximum number of features to
consider when looking for the best split. Later, we fit 20 trees into
a random forest method, which showed good performance but not a lot of
time waiting.

\section{Results}

In Figure \ref{vectorMSE}, we have the comparison of the individual
MSE per method, per response (18 functional markers). We can see in
this plot that the neural network model outperforms the other standard
approaches. In Figure \ref{vectorMSE}, we also see that the time is
better for the neural network approach.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figures/vector-mse.pdf}
\includegraphics[scale=0.35]{figures/time.pdf}
\caption{Vector MSE for all 18 predictors as in formula
  \ref{vectorMSEformula}, and running time in seconds.. We see that
  the performance of neural network is better than all the
  others, without being much slower.}
\label{vectorMSE}
\end{figure}

In Figure \ref{MSE}, we have the comparison of the joint MSE for all 18
responses. Here we can also see that the performance of neural network
is better than other approaches.

\begin{figure}
\centering
\includegraphics[scale=0.45]{figures/mse.pdf}
\caption{MSE for each 18 predictors as in formula
  \ref{MSEformula}. We see that the performance of neural
  network is better than all the others. Lines drawn only to
  appreciate the order among methods better.}
\label{MSE}
\end{figure}

\section{Discussion}

We showed that neural network is better than other approaches, mostly
due to the lack of simplifing assumptions like lienarity. The
performance is better in terms of prediction accuracy (measured by
mse) and computational time.

unlike regression, neural network takes advatange of the correlation
among responses by fitting a network for the whole response vector.

For future work, So, we have N experiments: N+1 matrices 100k by 50: B
(baseline), $E_1,...,E_N$. We want to use the baseline information to
predict the funcional markers at different experiments (see below for
more details) (also, surface markers do not change). That is, use B to
predict $E_i$ with a neural network. Using the individual 1 as
training, we can then only get baseline information for future
individuals and predict their functional markers. This is great
because this experiments are expensive, so our method would be widely
used to save money.

%\input{references} %%originally referenc

\end{document}

### 2018-09-25 script/exploratory.r

@@ -15,7 +15,7 @@ subdat = subset(dat1,select=c("X191.DNA","X193.DNA","X115.CD45","X139.CD45RA","X
 attr(subdat, "surface") <- c("X191.DNA","X193.DNA","X115.CD45","X139.CD45RA","X142.CD19","X144.CD11b","X145.CD4","X146.CD8","X148.CD34","X147.CD20","X158.CD33","X160.CD123","X167.CD38","X170.CD90","X110_114.CD3")
 attr(subdat, "functional") <- c("X141.pPLCgamma2","X150.pSTAT5","X152.Ki67","X154.pSHP2","X151.pERK1.2","X153.pMAPKAPK2","X156.pZAP70.Syk","X159.pSTAT3","X164.pSLP.76","X165.pNFkB","X166.IkBalpha","X168.pH3","X169.pP38","X171.pBtk.Itk","X172.pS6","X174.pSrcFK","X176.pCREB","X175.pCrkL")
 attributes(subdat)
-
+head(subdat)
 
 
 ## Histograms

### 2018-09-25 script/plots.r

@@ -1,0 +1,197 @@
+## R script for exploratory plots of raw data in data/cytof-5-data
+## Claudia April 2018
+
+datafolder = "../data/cytof-5-data/"
+dat1 = read.table(paste0(datafolder,"CyTOF54_Tube01_Day1_Unstim1_curated.fcs_eventnum_Ungated_Jnstim1_Day1_normalized_1_Unstim1_Singlets.fcs.txt"), header=TRUE, sep="\t")
+#dat2 = read.table(paste0(datafolder,"CyTOF54_Tube02_Day1_Unstim2_curated.fcs_eventnum_Ungated_Unstim2_Day1_normalized_2_Unstim2_Singlets.fcs.txt"), header=TRUE, sep="\t")
+#dat3 = read.table(paste0(datafolder,"CyTOF54_Tube03_Day2_Unstim3_curated.fcs_eventnum_Ungated_Unstim3_Day2_normalized_3_Unstim3_Singlets.fcs.txt"), header=TRUE, sep="\t")
+#dat4 = read.table(paste0(datafolder,"CyTOF54_Tube04_Day2_Unstim4_curated.fcs_eventnum_Ungated_Unstim4_Day2_normalized_4_Unstim4_Singlets.fcs.txt"), header=TRUE, sep="\t")
+#dat5 = read.table(paste0(datafolder,"CyTOF54_Tube05_Day3_Unstim5_curated.fcs_eventnum_Ungated_Unstim5_Day3_normalized_5_Unstim5_Singlets.fcs.txt"), header=TRUE, sep="\t")
+
+## extract only surface and functional markers
+subdat = subset(dat1,select=c("X191.DNA","X193.DNA","X115.CD45","X139.CD45RA","X142.CD19","X144.CD11b","X145.CD4","X146.CD8","X148.CD34","X147.CD20","X158.CD33","X160.CD123","X167.CD38","X170.CD90","X110_114.CD3","X141.pPLCgamma2","X150.pSTAT5","X152.Ki67","X154.pSHP2","X151.pERK1.2","X153.pMAPKAPK2","X156.pZAP70.Syk","X159.pSTAT3","X164.pSLP.76","X165.pNFkB","X166.IkBalpha","X168.pH3","X169.pP38","X171.pBtk.Itk","X172.pS6","X174.pSrcFK","X176.pCREB","X175.pCrkL"))
+
+## Identifying which ones are surface and which ones are functional
[…]

### 2018-09-26 manuscript/templates4Authors/latex_template/author.tex
first draft manuscript

@@
 \newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}
+\newcommand{\missing}[1]{\textcolor{red}{\textbf{#1}}}
 
@@
   Kundu, Peng Qiu, Daniel Pimentel-Alarc\'on}
-%\author{Xin Ma, Claudia Sol\'is-Lemus, Maxwell Hostetter II, Suprateek
-%  Kundu, Daniel Pimentel-Alarc\'on}
-%\author{xxx\thanks{work the same}, xxs$^*$}% \footnotemark[1]}
 % Use \authorrunning{Short Title} for an abbreviated version of
@@
 
-Despite the technological advances in acquiring an
-increasing number of parameters per single cell, approaches for
-analyzing such complex data lag behind. The existing approaches are
-often subjective and labor-intensive. For example, the widely used
-gating approach identifies cell types by user-defined sequences of
-nested 2-D plots. There have been efforts to develop clustering
-algorithms (eg, flowMeans, flowSOM, X-shift) and dimension reduction
-algorithms (eg, SPADE, tSNE, Scaffold). However, there is still huge
-space for developing new methods to ask new questions in this field.
+Despite the technological advances in acquiring an increasing number
+of parameters per single cell, approaches for analyzing such complex
+data lag behind. The existing approaches are often subjective and
+labor-intensive. For example, the widely used gating approach
+identifies cell types by user-defined sequences of nested 2-D
+plots. There have been efforts to develop clustering algorithms (e.g.,
+flowMeans\missing{(reference)}, flowSOM\missing{(reference)},
+X-shift\missing{(reference)}) and dimension reduction algorithms
+(e.g., SPADE\missing{(reference)}, tSNE\missing{(reference)},
+Scaffold\missing{(reference)}). However, there is still huge space
+for developing new methods to ask new questions in this field.
 % ---
@@
 interactions, among many others, by becoming a major discovery force
-in science due to the unprecedented accuracy in prediction.  Here, we
-explore neural network models to predict functional markers (internal
-phosphoproteins) with identify markers (cell surface proteins).
+in science due to the unprecedented accuracy in prediction.  Moreover,
+deep learning approaches have shown accurate performance on genomics
+and biomedical
+applications\cite{Ciresan2013,Denas2013,Fakoor2013,Leung2014,CruzRoa2014,Li2017,Mobadersany2018}.
 
+Furthermore, CyTOF data is perfectly suited for deep learning methods.
 On one side, identify markers define a cell type (e.g., B cell, T
@@
 functional markers, providing data for supervised learning tools, like
-neural networks.
+neural networks.  In addition, each CyTOF run typically collects data
+on $10^6$ cells, creating an ideal large dataset in which the number
+of samples (cells) is orders of magnitued larger than the number of
+variables (markers). Deep learning methods are particularly suited for
+this type of big data.
 
-\begin{itemize}
-\item Mass cytometry (CyTOF) is an emergent technological development
-  for high-dimensional multiparameter single cell analysis.
-\item CyTOF provides unprecedented multidimensional immune cell
-  profiling and has recently been applied to characterizing peripheral
-  blood cells, Natural Killer cells in viral infections, skin cells,
-  cells in celiac disease, responding phenotypes in cancer and even
-  holds the promise of examination of solid tumors (add references)
-\item Among the features per cell, there two classes: surface markers
-  (which identify the type of cell) and functional markers (which
-  identify the function of the cell)
-\item Collecting the functional markers is time-consuming and
-  computationally expensive (yes?), so it is desirable to devise a
-  prediction scheme that can allow us to predict the funcional markers
-  from the surface markers
-\item Deep learning methods have achieved outstanding performance in
-  various computational tasks, such as image analysis, natural
-  language processing, and pattern recognition (references)
-\item Moreover, recent efforts to use deep learning approaches in
-  genomics and biomedical applications show their flexibility for
-  handling complex problems (Cires ̧an et al., 2013; Cruz-Roa et al.,
-  2014; Denas and Taylor, 2013; Fakoor et al., 2013; Leung et al.,
-  2014).
-\item in each FCM and CyTOF run we typically collect approximately $10^5$
-  to $10^6$ cells, so that the number of instances (cells) is several
-  orders of magnitude larger than the number of variables (up to 50
-  markers). Therefore, developing deep learning approaches to analyze
-  cytometry data is very promising.
-\item We present DeepCyTOF, an integrated deep learning domain
-  adaptation framework, which employs one manually gated reference
-  sample and utilizes it for automated gating of the remaining samples
-  in a study.
-\item We use this method in the dataset bla
-\item We compare the NN approach to other usual prediction methods
-  like regression-based and xxx (Suprateek is checking this), and show
-  how the NN approach improves the prediction by the incorporation of
-  non-linear and interaction effects.
-\item We show the improvement driven by NN when the effects are
-  nonlinear as a powerful alternative to the usual regression techniques.
-\end{itemize}
+Here, we explore neural network models to predict functional markers
+(internal phosphoproteins) with identify markers (cell surface
+proteins), and compare its performance in terms of accuracy and speed
+to other standard statistical approaches like regression, and random
+forests.  We show that neural networks improve prediction of
+functional markers, making them a powerful alternative to the usual
+regression techniques.
 
@@
 
-In a dataset we previously published
-[Bendall et al 2011; Qiu et al 2011], we measured 13 identity
-(surface) markers and 18 functional markers simultaneously.  This
-measurement was performed on 5 health donor bone marrow samples
-untreated, and also on 24 health bone marrow samples treated by 24
-different drugs.  For the moment, let’s focus on the first 5 samples.
+\subsection{Pre-processing}
 
-Given the data from 1 of the 5 samples, we have observations (data
-points) for roughly 250,000 cells. This 250,000*(13+18) data can be
-used to build a machine learning model to predict the 18 function
-markers based on the 13 identity markers. In this dataset, we roughly
-know that the 13 identity markers can be used to define ~25 cell
-types, and we have manually created cell type label for each cell. One
-machine learning question can be predicting the cell type label based
-on the 13 surface markers, however, this is not the question we are
-asking here. The questions we want to ask are:
+The dataset has been previously published in
+\cite{Bendall2011,Qiu2011}. It contains measurements for 13 identity
+(surface) markers, and 18 functional markers, which were performed on
+5 healthy untreated donor bone marrow samples. The data roughly
+contains 250,000 cells per sample (1,223,228 cells in
+total). Thus, the data can be expressed in a $1223228 \times 31$
+matrix.
 
-1. If we use 1 sample
-to train a deep learning model which use identity markers to predict
-functional markers, can this model be used (generalized) to predict
-functional markers in the other 4 untreated samples?
+The data was transformed with inverse hyperbolic sine function
+($arcsinh$ with co-factor of 5), which is the standard transformation
+for CyTOF data\missing{(reference)}.
 
-2. Based on the
-trained deep learning model, can we try to identify cell clusters? And
-will our cell clusters agree with the well-accepted cell types in the
-literature.
+\missing{(Peng: what was done to the data prior to sharing with us?)}
 
-3. If we focus cells belonging to the same known cell
-type, and examine the distribution of functional markers and the
-correlation with the subtle variations of the identity markers among
-cells of this type, are we able to see hints that this cell type may
-be further divided into subtypes?
+\subsection{Exploratory analysis}
 
-The data is highly complex, violating the linearity assumptions of
-linear regression, see plot, even after the standard transformation
-for CyTOF data (arcsinh with cofactor=5). \ref{exploratory}
+We will compare the performance different methods (explained next
+section) to predict the functional markers with the surface markers.
+The data is highly complex and correlated, violating some of the
+fundamental assumptions of standard statistical approaches (like
+regression). For example, the data is highly skewed and the pattern
+between response and predictor is not linear (see figure
+\ref{exploratory}).
 
@@
   histograms show a biased pattern, and the scatterplot shows non
-  linearity, crucial assumptions in regression.}
+  linearity, both violations of crucial assumptions in standard
+  regression models.}
 \label{exploratory}
@@
 
+\subsection{Background on neural networks}
+
+A neural network model is formed by several {\em neurons}. Each neuron
+receives an input vector $x$, then weights its components according to
+the neuron's weight vector $w$, adds a {\em bias} constant $b$, and
+passes the result through a non-linear {\em activation function}
+$\sigma$. This way, the output of a neuron is given by $\sigma(w^Tx +
+b)$. There are several options for the activation function
+$\sigma$. Common choices include the {\em sigmoid} function
+$\sigma(z)=\frac{1}{1+e^-z}$ or the {\em rectified linear unit} (ReLU)
+$\sigma(z)=\max(0,z)$. For the CyTOF data, we use the hyperbolic
+tangent as activation function, as it showed better performance than
+the sigmoid or ReLU functions (more details on the specific neural
+network fit in subsection \ref{comparison}).
+
+The final output of the network is given by $\hat{f}(x)$ with
+parameters $\mathbf{W}_1,\cdots,\mathbf{W}_L$ for the weight matrices
+and $b_1,\cdots,b_L$ for the bias vectors for each layer.
+
+The estimation of the parameters is done through the following
+optimization
+\begin{equation}
+\min_{\{W_l,b_l\}_{l=1}^L} \ \ \sum_{i=1}^n \| y_i-\hat{f}(x_i) \|^2.
+\end{equation}
+
+The most widely used technique to solve this optimization is through
+stochastic gradient descent (SGD) and back-propagation, but we
+discovered that Adam\cite{Kingma2014}, an algorithm for first-order
+gradient-based optimization of stochastic objective functions, based
+on adaptive estimates of lower-order moments had better performance
+for our data (more details in subsection \ref{comparison}).
+
+\subsection{Methods comparison}
+\label{comparison}
+
 We fit a neural network model to predict functional markers from
 surface markers, and compare its performance to three classical
-statistical methods like linear regression (unpenalized and
-penalized), and random trees. Due to computational time constraints,
-we could not fit a support vector regression in the dataset. We
-compared the performance of the four approaches by computing the mean
-square error of the predicted responses.
+statistical methods: 1) linear regression (unpenalized and penalized),
+2) decision trees, and 3) random trees. Due to computational time
+constraints, we could not fit a support vector regression (SVR)
+model. We compared the performance of the four approaches by computing
+the mean square error (MSE) of the predicted responses.
 
-To fit the models, we divided the data (1,223,228 cells in 18
-functional markers (responses) and 15 surface markers (predictors))
-into three subsets: 250,000 rows as training data, 250,000 rows as
-validation data and 223,228 rows as testing data. The training data
-was used to fit each of the four models. The validation data was used
-to determine the best setup (tuning parameters) of each model in terms
-of mse, and the test data was used to compare the mse across methods.
+To fit the models, we divided the data into training set, validation
+set and testing set. The training set was used to fit each of the four
+models. The validation set was used to determine the best setup
+(tuning parameters) of each model in terms of MSE. Finally, the test
+set was used to compare the MSE across methods.
 
-We used the vector mse (put formula), and the 18 individual mse's (put
-formula) on the test data to compare across methods.
+The complete data consisted of 1,223,228 cells in 18 functional
+markers (responses) and 15 surface markers (predictors), which we
+divided as follows: 250,000 rows as training set, 250,000 rows as
+validation set, and 223,228 rows as testing set.
 
+We used two separate measure of performance: a vector MSE (equation
+\ref{vectorMSEformula}) and individual MSE (equation
+\ref{MSEformula}), one per predictor (so, 18 in total).
+
+The vector MSE is defined as
 \begin{equation}
@@
 \end{equation}
-
 where $\hat{Y}_i \in \mathbf{R}^{18}$ is the predicted vector of
@@
 
-The individual MSE was then defined for predictor $k$ as
+The individual MSE for predictor $k$ is defined  as
 \begin{equation}
@@
 \vspace{0.25cm}
-\noindent \textbf{Neural network model} We tested different network
+\noindent \textbf{Neural network model:} We tested different network
 architectures, activation functions, regularization coefficients,
 solver methods, momentum policies, and learning rates. The best
-network has four layers (see figure xxx) with 90, 90, 45 and 45
+network has four layers (see figure \ref{nn}) with 90, 90, 45 and 45
 nodes. The network uses hyperbolic tangent as activation function,
-regularization coefficient 0.0001, momentum policy at 0.8, learning
-rates of (0.05, 0.01, 0.0001, 0.75).
+regularization coefficient of 0.0001, momentum policy at 0.8, learning
+rates of 0.01, 0.0001, 0.75. We used Adam solver\cite{Kingma2014}, an
+algorithm for first-order gradient-based optimization of stochastic
+objective functions, based on adaptive estimates of lower-order
+moments, instead of Stochastic Gradient Descent (SGD) as the former
+showed increased accuracy. All networks were trained using the julia
+package Mocha\missing{(reference)}.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.4]{figures/neuralNet.eps}
+\caption{Neural network for predicting functional markers (18
+  responses) from surface markers (15 predictors) with 4 hidden
+  layers with 90, 90, 45 and 45 nodes each.}
+\label{nn}
+\end{figure}
 
 \vspace{0.25cm}
-\noindent \textbf{Linear regression (unpenalized/penalized)} We fit standard
-linear regression, as well as the penalized version with LASSO penalty
-under different penalization parameters. We used the
-\textit{ScikitLearn} julia wrapper, with default settings.
+\noindent \textbf{Linear regression (unpenalized/penalized):} We fit
+standard linear regression, as well as the penalized version with
+LASSO penalty under different penalization parameters. We used the
+\textit{ScikitLearn}\missing{(reference)} julia wrapper, with default
+settings. We noted that the penalized version performed worse than the
+unpenalized version for all the predictors (regardless of penalty
+parameter), so we only present results below for the unpenalized
+linear regression model.
 
 \vspace{0.25cm}
-\noindent \textbf{Decision tree and random trees} We fit one decision
+\noindent \textbf{Decision tree and random trees:} We fit one decision
 regression tree with \textit{ScikitLearn} julia wrapper, with default
@@
 is the default setting. We did not constraint the maximum depth of the
-tree, and used as 2 the minimum number of samples required to split an
-internal node. We did not constraint the maximum number of features to
-consider when looking for the best split. Later, we fit 20 trees into
-a random forest method, which showed good performance but not a lot of
-time waiting.
+tree, and set as 2 the minimum number of samples required to split an
+internal node. In addition, we did not constraint the maximum number
+of features to consider when looking for the best split. Later, we fit
+20 trees into a random forest method. We could explore more than 20
+trees due to computational time constraints.
 
@@
 
-In Figure \ref{vectorMSE}, we have the comparison of the individual
-MSE per method, per response (18 functional markers). We can see in
-this plot that the neural network model outperforms the other standard
-approaches. In Figure \ref{vectorMSE}, we also see that the time is
-better for the neural network approach.
+Figure \ref{vectorMSE} (left) shows the vector MSE (equation
+\ref{vectorMSEformula}) across all four different methods, being
+decision tree the least accurate and neural network the most
+accurate. Figure \ref{vectorMSE} (right) shows a comparison on
+computation time (in seconds) among the four methods, being linear
+regression the fastest and random forest the slowest. To sum up, the
+neural network approach outperforms the other three methods in terms
+of prediction accuracy, without sacrificing too much computational
+speed.
 
@@
 \includegraphics[scale=0.35]{figures/time.pdf}
-\caption{Vector MSE for all 18 predictors as in formula
-  \ref{vectorMSEformula}, and running time in seconds.. We see that
-  the performance of neural network is better than all the
-  others, without being much slower.}
+\caption{Left: Vector MSE (equation \ref{vectorMSEformula}) for all
+  four methods sorted from most accurate (neural network) to least
+  accurate (decision tree). Right: Running time (in seconds) for the
+  training and validation sets (sample 750,000 rows) for all four
+  methods, sorted from fastest (linear regression) to slowest (random
+  forest).}
 \label{vectorMSE}
@@
 
-In Figure \ref{MSE}, we have the comparison of the joint MSE for all 18
-responses. Here we can also see that the performance of neural network
-is better than other approaches.
+Figure \ref{MSE} shows the individual MSE (equation \ref{MSEformula})
+per predictor (18 predictors in x-axis) for each of the four
+methods. Again, the prediction accuracy of neural network is better
+than the other three methods for all the 18 predictors.
 
@@
 \includegraphics[scale=0.45]{figures/mse.pdf}
-\caption{MSE for each 18 predictors as in formula
-  \ref{MSEformula}. We see that the performance of neural
-  network is better than all the others. Lines drawn only to
-  appreciate the order among methods better.}
+\caption{Individual MSE (equation \ref{MSEformula}) for each 18
+  predictors. The neural network outperforms all other methods across
+  all predictors.  Lines are drawn simply for visual effect.}
 \label{MSE}
+\end{figure}
+
+Finally, we present a figure of the predicted response vs the observed
+response, for all the four methods and 18 functional markers (figure
+\ref{predObs}).
+
+\begin{figure}
+\centering
+%\includegraphics[scale=0.4]{figures/neuralNet.eps}
+\caption{Predicted vs observed responses across all functional markers
+  and all four methods. The closer the slope to 1, the
+  better.\missing{(missing this plot)}.}
+\label{predObs}
 \end{figure}
@@
 
-We showed that neural network is better than other approaches, mostly
-due to the lack of simplifing assumptions like lienarity. The
-performance is better in terms of prediction accuracy (measured by
-mse) and computational time.
+In this work, we showed that a neural network model outperforms
+standard statistical approaches like linear regression and random
+forest in the prediction of functional markers from surface markers
+for CyTOF data. Neural networks were also faster and more efficient
+than random forests, which make them a more viable choice for big
+datasets.
 
-unlike regression, neural network takes advatange of the correlation
+The improved prediction accuracy of neural networks can be explained
+by their flexibility to account for non-linearity or skewness.  Unlike
+regression models, neural networks do not have linearity or normality
+assumptions, and they take advatange of the correlation structure
 among responses by fitting a network for the whole response vector.
 
-For future work, So, we have N experiments: N+1 matrices 100k by 50: B
-(baseline), $E_1,...,E_N$. We want to use the baseline information to
-predict the funcional markers at different experiments (see below for
-more details) (also, surface markers do not change). That is, use B to
-predict $E_i$ with a neural network. Using the individual 1 as
-training, we can then only get baseline information for future
-individuals and predict their functional markers. This is great
-because this experiments are expensive, so our method would be widely
-used to save money.
+As mentioned already, CyTOF data is perfectly suited for deep learning
+methods given the simulteanous measurement of a large number of
+protein markers, including both both identity markers and functional
+markers. Both measurements allow for the implementation of highly
+accurate supervised methods, like neural networks. In addition, the
+structure of CyTOF data is ideal for deep learning: number of samples
+orders of magnited greater than the number of variables.
+
+For future work, we can include an extended version of the
+dataset\cite{Bendall2011,Qiu2011} that includes 24 healthy sample of
+bone marrow treated by 24 different drugs. In this setting, we are
+interested in predicting the functional markers under different drug
+scenarios, using information at baseline (no treatment) and surface
+markers at different treatment levels.
+Furthermore, based on the trained deep learning model, we are
+interested in the question of whether we can identify cell clusters,
+and whether these cell clusters agree with well-accepted cell types in
+literature.
+Finally, if we focus on cells belonging to the same known cell type, and
+examine the distribution of functional markers and the correlation
+with the subtle variations of the identity markers among cells of this
+type, we can explore whether there is evidence that the specific cell
+type could be further divided into subtypes.
+
+% For future work, So, we have N experiments: N+1 matrices 100k by 50: B
+% (baseline), $E_1,...,E_N$. We want to use the baseline information to
+% predict the funcional markers at different experiments (see below for
+% more details) (also, surface markers do not change). That is, use B to
+% predict $E_i$ with a neural network. Using the individual 1 as
+% training, we can then only get baseline information for future
+% individuals and predict their functional markers. This is great
+% because this experiments are expensive, so our method would be widely
+% used to save money.
 
 %\input{references} %%originally referenc
+\bibliography{neural-nets}
+\bibliographystyle{nihunsrt} % Use the custom nihunsrt bibliography style included with the template
 

### 2018-11-14 manuscript/arxiv/nn-cytof.tex
finished paper

--- 2018-09-26 manuscript/templates4Authors/latex_template/author.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample root file for your "contribution" to a contributed volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[conference]{IEEEtran}
 
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[graybox]{svmult}
+%========================================================================= PACKAGES
+\usepackage{amsmath,amssymb}
+\usepackage[mathscr]{euscript} %To be able to use script letters with \mathscr{}
+\usepackage{amsthm} %To be able to write Definitions, Theorems, etc.
+\usepackage{tikz} %To draw graphs.
+\usepackage{graphicx} % To scale equations
+\newcommand*{\Scale}[2][4]{\scalebox{#1}{\ensuremath{#2}}} % To scale equations
+\usepackage{framed} % To be able to frame theorems and stuff
+%\usepackage{float} %To be able to place figures exactly where I want with [H]
+\usepackage{multirow}
+\usepackage{color}
+\usepackage{subfigure}
+\usepackage{longtable}
+\usepackage{MnSymbol}
+\usepackage{xcolor}  \definecolor{shadecolor}{rgb}{.95,.95,.95}  %To put a shaded region
+\usepackage[numbers,square, comma, sort&compress]{natbib}	 %Make the references [1-8], rather than [1]-[8], as with \usepackage{cite}
+\renewcommand*{\bibfont}{\footnotesize} %Size of the references; only needed with natbib
+%\usepackage{makeidx} %To create an index
+%\makeindex
+\usepackage[hidelinks]{hyperref} %To be able to use links inside the document; [hidelinks removes the ugly red boxes]
 
-% choose options for [] as required from the list
-% in the Reference Guide
+%==================================================== ALGORITHM SETUP
+\usepackage[ruled,vlined]{algorithm2e}
 
-\usepackage{mathptmx}       % selects Times Roman as basic font
-\usepackage{helvet}         % selects Helvetica as sans-serif font
-\usepackage{courier}        % selects Courier as typewriter font
-\usepackage{type1cm}        % activate if the above 3 fonts are
-                            % not available on your system
-%
-\usepackage{makeidx}         % allows index generation
-\usepackage{graphicx}        % standard LaTeX graphics tool
-                             % when including figure files
-\usepackage{multicol}        % used for the two-column index
-\usepackage[bottom]{footmisc}% places footnotes at page bottom
+%========================================================================= THEOREMS, LEMMAS, DEFINITIONS, ETC.
+\newtheorem{myDefinition}{Definition}
+\newtheorem{myTheorem}{Theorem}
+\newtheorem{myLemma}{Lemma}
+\newtheorem{myCorollary}{Corollary}
+\newtheorem{myProposition}{Proposition}
+\newtheorem{myExample}{Example}
+\newtheorem{myRemark}{Remark}
+\newtheorem{myConjecture}{Conjecture}
+\newtheorem{myQuestion}{Question}
+\newtheorem{myProblem}{Problem}
 
-% see the list of further useful packages
-% in the Reference Guide
+%========================================================================= STUFF TO DRAW GRAPHS
+\tikzstyle{every edge}=  [draw]
+\tikzstyle{vertex} = [draw,circle,minimum size=1pt]
+\tikzstyle{label} = [minimum size=.1pt,font=\scriptsize]
+\tikzstyle{title} = [minimum size=.25cm,font=\small]
 
-\makeindex             % used for the subject index
-                       % please use the style svind.ist with
-                       % your makeindex program
+%========================================================================= SYMBOLS
+%\newcommand{\var}[1]{\mathsf{#1}}
+\newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\careful}{\textcolor{red}}
+\newcommand{\comment}{\textcolor{blue}}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Miscelaneous
+%\def \R{\mathbb{R}}
+%\def \P{\mathsf{P}}
+%\def \T{\mathsf{T}}
+%\def \c{\mathsf{c}}
+\def \spn{{\rm span}}
+\def \rnk{{\rm rank}}
 
-%% track changes
-\usepackage[normalem]{ulem}
-\usepackage[dvipsnames]{xcolor}
-\newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}
-\newcommand{\missing}[1]{\textcolor{red}{\textbf{#1}}}
+%Scalars
+\def \r{{\hyperref[rDef]{r}}}
+
+%text
+\def \as{{\hyperref[asDef]{{\rm a.s.}}}}
+%Assumptions
+\def \Cone{{\hyperref[ConeDef]{{\rm {\bf C1}}}}}
+%Theorems
+\def \mainThm{{\hyperref[mainThm]{Theorem \ref{mainThm}}}}
+%Sections
+\def \modelSec{{\hyperref[modelSec]{Section \ref{modelSec}}}}
+%Figures
+\def \distributionFig{{\hyperref[distributionFig]{Figure \ref{distributionFig}}}}
+%Algorithms
+\def \validationAlg{{\hyperref[validationAlg]{Algorithm \ref{validationAlg}}}}
+
+\newcounter{@inst}
+\newcounter{@auth}
+\newcounter{auco}
+\newdimen\instindent
+\newbox\authrun
+\newtoks\authorrunning
+\newtoks\tocauthor
+\newbox\titrun
+\newtoks\titlerunning
+\newtoks\toctitle
+
+\def\clearheadinfo{\gdef\@author{No Author Given}%
+                   \gdef\@title{No Title Given}%
+                   \gdef\@subtitle{}%
+                   \gdef\@institute{}%
+                   \gdef\@thanks{}%
+                   \global\titlerunning={}\global\authorrunning={}%
+                   \global\toctitle={}\global\tocauthor={}}
+
+\def\institute#1{\gdef\@institute{#1}}
+
+\def\institutename
+ {\begingroup
+ \if!\@institute!\else
+ \def\thanks##1{\unskip{}}%
+ \def\iand{\\[5pt]\let\and=\nand}%
+ \def\nand{\ifhmode\unskip\nobreak\fi\ $\cdot$ }%
+ \let\and=\nand
+ \def\at{\\\let\and=\iand}%
+ \footnotetext[0]{\kern-\bibindent
+ \ignorespaces\@institute}\vspace{5dd}\fi
+ \endgroup
+ }%
+
+% correct bad hyphenation here
+%\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 
-\title*{Prediction of functional markers of mass cytometry data via
+\title{Prediction of functional markers of mass cytometry data via
   deep learning}
@@
 \email{suprateek.kundu@emory.edu}
-\and Peng Qiu \at Georgia Institute of Technology, Atlanta, GA,
+\and Peng Qiu \at Georgia Institute of Technology and Emory University, Atlanta, GA,
 \email{peng.qiu@bme.gatech.edu}
@@
 
-\abstract{add here}
+\begin{abstract}
+Recently, there has been an increasing interest in the analysis of
+flow cytometry data, which involves measurements of a set of surface
+and functional markers across hundreds and thousands of cells. These
+measurements can often be used to differentiate various cell types and
+there has been a rapid development of analytic approaches for
+achieving this. However, in spite of the fact that measurements are
+available on such a large number of cells, there have been very
+limited advances in deep learning approaches for the analysis of flow
+cytometry data. Some preliminary work has focused on using deep
+learning techniques to classify cell types based on the cell protein
+measurements. In a first of its' kind study, we propose a novel deep
+learning architecture for predicting functional markers in the cells
+given data on surface markers. Such an approach is expected to
+automate the measurement of functional markers across cell samples,
+provided data on the surface markers are available, that has important
+practical advantages. We validate and compare our approach with
+competing machine learning methods using a real flow cytometry
+dataset, and showcase the improved prediction performance of the deep
+learning architecture.
+\end{abstract}
 
@@
 plots. There have been efforts to develop clustering algorithms (e.g.,
-flowMeans\missing{(reference)}, flowSOM\missing{(reference)},
-X-shift\missing{(reference)}) and dimension reduction algorithms
-(e.g., SPADE\missing{(reference)}, tSNE\missing{(reference)},
-Scaffold\missing{(reference)}). However, there is still huge space
+flowMeans\cite{Aghaeepour2010}, flowSOM\cite{VanGassen2015},
+X-shift\cite{Samusik2016}, and dimension reduction algorithms
+(e.g., SPADE \cite{Qiu2011}, tSNE\cite{vanderMaaten2008},
+Scaffold\cite{Spitzer2015}. However, there is still huge space
 for developing new methods to ask new questions in this field.
@@
 secreting cytokines, proliferating, apoptosis). Since CyTOF technology
-allows for the simulteanous measurement of a large number of protein
+allows for the simultaneous measurement of a large number of protein
 markers, most CyTOF studies measure both identity markers and
@@
 on $10^6$ cells, creating an ideal large dataset in which the number
-of samples (cells) is orders of magnitued larger than the number of
+of samples (cells) is orders of magnitude larger than the number of
 variables (markers). Deep learning methods are particularly suited for
 this type of big data.
+
+In terms of motivation, there are two main reasons to predict the functional markers from surface markers in CyTOF data: 1) monetary and time cost, and 2) technical limit of the total number of markers CyTOF can measure, which is currently around 50 protein markers.
+That is, if we can accurately predict some functional markers based on surface markers, there is no longer the need to include those functional markers in the staining panel (experimental design), and thereby freeing up channels to measure more surface markers or additional functional markers that cannot be predicted.
 
@@
 
+\begin{figure*}
+\centering
+\includegraphics[scale=0.35]{figures/surface-hist.pdf}
+\includegraphics[scale=0.35]{figures/functional-hist.pdf}
+\includegraphics[scale=0.35]{figures/scatterplot.pdf}
+\caption{Exploratory plots of surface and functional markers. The
+  histograms show a biased pattern, and the scatterplot shows non
+  linearity, both violations of crucial assumptions in standard
+  regression models.}
+\label{exploratory}
+\end{figure*}
+
 \subsection{Pre-processing}
 
-The dataset has been previously published in
-\cite{Bendall2011,Qiu2011}. It contains measurements for 13 identity
-(surface) markers, and 18 functional markers, which were performed on
-5 healthy untreated donor bone marrow samples. The data roughly
-contains 250,000 cells per sample (1,223,228 cells in
-total). Thus, the data can be expressed in a $1223228 \times 31$
+The CyTOF dataset has been previously published in
+\cite{Bendall2011,Qiu2011}.
+It contains single-cell data for 5 bone marrow samples from healthy donors.
+The data for each sample contains measurements for 31 protein markers for individual cells, including 13 cell surface markers which are conventionally used to define cell types, as well as 18 functional markers which reflect the signaling activities of cells.
+The number of cells per sample is roughly 250,000, and the total number of cells across all 5 samples is 1,223,228.
+Thus, the data can be expressed in a $1223228 \times 31$
 matrix.
@@
 ($arcsinh$ with co-factor of 5), which is the standard transformation
-for CyTOF data\missing{(reference)}.
-
-\missing{(Peng: what was done to the data prior to sharing with us?)}
+for CyTOF data \cite{Bendall2011}.
 
@@
 
-We will compare the performance different methods (explained next
+We will compare the performance of different methods (explained next
 section) to predict the functional markers with the surface markers.
@@
 \ref{exploratory}).
-
-\begin{figure}
-\includegraphics[scale=0.3]{figures/surface-hist.pdf}
-\includegraphics[scale=0.3]{figures/functional-hist.pdf}
-\includegraphics[scale=0.3]{figures/scatterplot.pdf}
-\caption{Exploratory plots of surface and functional markers. The
-  histograms show a biased pattern, and the scatterplot shows non
-  linearity, both violations of crucial assumptions in standard
-  regression models.}
-\label{exploratory}
-\end{figure}
 
@@
 markers (responses) and 15 surface markers (predictors), which we
-divided as follows: 250,000 rows as training set, 250,000 rows as
+divided as follows: 750,000 rows as training set, 250,000 rows as
 validation set, and 223,228 rows as testing set.
@@
 \begin{equation}
-MSE_{vec} = \frac{1}{n} \sum_{i=1}^n ||\hat{Y}_i - Y_i ||^2_2
+MSE_{vec} = \frac{1}{2n} \sum_{i=1}^n ||\hat{Y}_i - Y_i ||^2_2
 \label{vectorMSEformula}
@@
 \begin{equation}
-MSE_{(k)} = \frac{1}{n} \sum_{i=1}^n (\hat{Y}_{k,i} - Y_{k,i} )^2
+MSE_{(k)} = \frac{1}{2n} \sum_{i=1}^n (\hat{Y}_{k,i} - Y_{k,i} )^2
 \label{MSEformula}
@@
 
-% https://github.com/pluskid/Mocha.jl/blob/master/docs/user-guide/solver.rst
-\vspace{0.25cm}
-\noindent \textbf{Neural network model:} We tested different network
-architectures, activation functions, regularization coefficients,
-solver methods, momentum policies, and learning rates. The best
-network has four layers (see figure \ref{nn}) with 90, 90, 45 and 45
-nodes. The network uses hyperbolic tangent as activation function,
-regularization coefficient of 0.0001, momentum policy at 0.8, learning
-rates of 0.01, 0.0001, 0.75. We used Adam solver\cite{Kingma2014}, an
-algorithm for first-order gradient-based optimization of stochastic
-objective functions, based on adaptive estimates of lower-order
-moments, instead of Stochastic Gradient Descent (SGD) as the former
-showed increased accuracy. All networks were trained using the julia
-package Mocha\missing{(reference)}.
-
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \label{nn}
-\end{figure}
+\end{figure*}
+
+% https://github.com/pluskid/Mocha.jl/blob/master/docs/user-guide/solver.rst
+\vspace{0.25cm}
+\noindent \textbf{Neural network model:} We tested different network
+architectures, activation functions, regularization coefficients,
+solver methods, momentum policies, and learning rates with 50,000 maximum epochs. The best
+network has four layers (see figure \ref{nn}) with 90, 90, 45 and 45
+nodes. The network uses hyperbolic tangent as activation function,
+regularization coefficient of 0.0001, momentum policy fixed at 0.8, inverse-decay learning
+rate policy with base learning rate, gamma and power parameters at 0.01, 0.0001, 0.75. We used Adam solver\cite{Kingma2014}, an
+algorithm for first-order gradient-based optimization of stochastic
+objective functions, based on adaptive estimates of lower-order
+moments, instead of Stochastic Gradient Descent (SGD) as the former
+showed increased accuracy. All networks were trained using the julia
+package \texttt{Mocha}\cite{Bezanson2017,mocha}.
 
@@
 LASSO penalty under different penalization parameters. We used the
-\textit{ScikitLearn}\missing{(reference)} julia wrapper, with default
+\texttt{ScikitLearn}\cite{scikit} julia wrapper, with default
 settings. We noted that the penalized version performed worse than the
@@
 \vspace{0.25cm}
-\noindent \textbf{Decision tree and random trees:} We fit one decision
-regression tree with \textit{ScikitLearn} julia wrapper, with default
+\noindent \textbf{Decision tree and random forest regressions:} We fit one decision tree
+regression per response with \texttt{ScikitLearn} julia wrapper, with default
 settings. We compared the performance of the ``mse'' criterion and the
@@
 of features to consider when looking for the best split. Later, we fit
-20 trees into a random forest method. We could explore more than 20
+20 trees into a random forest regression. We could not explore more than 20
 trees due to computational time constraints.
@@
 \centering
-\includegraphics[scale=0.35]{figures/vector-mse.pdf}
-\includegraphics[scale=0.35]{figures/time.pdf}
+\includegraphics[scale=0.45]{figures/vector-mse.pdf}
+\includegraphics[scale=0.45]{figures/time.pdf}
 \caption{Left: Vector MSE (equation \ref{vectorMSEformula}) for all
@@
 Figure \ref{MSE} shows the individual MSE (equation \ref{MSEformula})
-per predictor (18 predictors in x-axis) for each of the four
+per response (18 responses in x-axis) for each of the four
 methods. Again, the prediction accuracy of neural network is better
@@
 \caption{Individual MSE (equation \ref{MSEformula}) for each 18
-  predictors. The neural network outperforms all other methods across
-  all predictors.  Lines are drawn simply for visual effect.}
+  responses. The neural network outperforms all other methods across
+  all responses.  Lines are drawn simply for visual effect.}
 \label{MSE}
@@
 
-Finally, we present a figure of the predicted response vs the observed
-response, for all the four methods and 18 functional markers (figure
-\ref{predObs}).
+The MSE performance varies across responses.
+For example, the first response (functional marker 141.pPLCgamma2) has an overall MSE lower than other responses like the third (functional marker 152.Ki67), the 8th (functional marker 159.pSTAT3) or the 14th (functional marker 171.pBtk.Itk).
 
-\begin{figure}
+Figure \ref{violin} (left) shows the violin plots for these 4 functional markers.
+We observe that the 14th response has a wider range and heavier tails than the other responses, which is confirmed in the scatterplots on the center and right (figure \ref{violin}).
+It appears that the wider spread and higher variability of the 14th response (functional marker 171.pBtk.Itk) causes the lower prediction accuracy compared to other responses, like the first one (functional marker 141.pPLCgamma2).
+
+\begin{figure*}
 \centering
-%\includegraphics[scale=0.4]{figures/neuralNet.eps}
-\caption{Predicted vs observed responses across all functional markers
-  and all four methods. The closer the slope to 1, the
-  better.\missing{(missing this plot)}.}
+\includegraphics[scale=0.35]{figures/violin.pdf} \hspace{0.25cm}
+\includegraphics[scale=0.35]{figures/response1.pdf}
+\hspace{0.25cm}
+\includegraphics[scale=0.35]{figures/response14.pdf}
+\caption{Left: Violin plot for four functional markers (responses).
+Horizontal lines represent the 25th quantile, median and 75th quantile.
+Center: Predicted vs observed responses on the first functional marker (141.pPLCgamma2)
+  across all four methods. Left: Predicted vs observed responses on the 14th functional marker (171.pBtk.Itk)
+  across all four methods. The closer the slope to 1 (black line), the
+  better.}
+\label{violin}
+\end{figure*}
+
+Finally, we present selected scatterplots of surface markers as predictors for the responses 1,3,8 and 14 (figure
+\ref{predObs}). We can appreciate in these plots the non-linear relationship between the predictors and responses, which justifies the use of a neural network approach.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.3]{figures/scatterplot1.pdf} \hspace{0.5cm}
+\includegraphics[scale=0.3]{figures/scatterplot14.pdf}
+\caption{Scatterplot of selected surface markers (predictors) and selected functional markers (responses).
+Left: the first response (functional marker 141.pPLCgamma2) shows a linear relationship to the predictor (surface marker 115.CD45), which partially explains the better MSE in figure \ref{MSE}.
+Right: the 14th response (functional marker 171.pBtk.Itk) shows a non-linear relationship to the predictor (surface marker 115.CD45), which partially explains the worse MSE in figure \ref{MSE}.}
 \label{predObs}
@@
 regression models, neural networks do not have linearity or normality
-assumptions, and they take advatange of the correlation structure
+assumptions, and they take advantage of the correlation structure
 among responses by fitting a network for the whole response vector.
@@
 As mentioned already, CyTOF data is perfectly suited for deep learning
-methods given the simulteanous measurement of a large number of
-protein markers, including both both identity markers and functional
+methods given the simultaneous measurement of a large number of
+protein markers, including both identity markers and functional
 markers. Both measurements allow for the implementation of highly
@@
 structure of CyTOF data is ideal for deep learning: number of samples
-orders of magnited greater than the number of variables.
+orders of magnitude greater than the number of variables.
+
+The accuracy in the prediction of functional markers from surface markers has economic and computational advantages, for example, considering the limitation to the total number of markers CyTOF can measure, which is currently around 50 protein markers.
+Being able to predict functional markers from surface markers could allow for different types of staining panels which could measure more surface markers, or focus on functional markers not so easily predicted.
 
@@
 
+% %======================================================= MAIN THEOREM
+% \begin{framed}
+% \begin{myTheorem}
+% \label{mainThm}
+% Write Theorem
+% \end{myTheorem}
+% \end{framed}
+
 \end{document}

### 2018-11-14 script/plots.r

@@ -181,12 +181,12 @@ p <- ggplot(dff,aes(x=pred,y=mse,col=method))+geom_point() + geom_line()+
         axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
         axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
         axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        legend.text=element_text(size=rel(1.2)), legend.title=element_text(size=rel(1.5)),
-        panel.background = element_blank(),
+        legend.text=element_text(size=rel(1.2)),
+        panel.background = element_blank(),legend.title=element_blank(),
         axis.line = element_line(colour = "grey")
         ) +
     scale_x_continuous(breaks=c(1:18))+
-    xlab("Predictor") + ylab("MSE")
+    xlab("Response") + ylab("MSE")
 
 
 pdf("mse.pdf",width=8,height=5.5)
@@ -195,3 +195,179 @@ dev.off()
 
 
 
+########
+## Plots per response
+## We will focus in response 1 (good), 3, 8, 14 (bad)
+
+dat1 = read.table("../results/Predictions/Test_1.txt", header=TRUE)
+dat3 = read.table("../results/Predictions/Test_3.txt", header=TRUE)
+dat8 = read.table("../results/Predictions/Test_8.txt", header=TRUE)
+dat14 = read.table("../results/Predictions/Test_14.txt", header=TRUE)
+
+newdat = data.frame(response = c(dat1$True,dat3$True,dat8$True,dat14$True),
+    which = c(rep(1,length(dat1$True)),rep(3,length(dat3$True)),rep(8,length(dat8$True)),rep(14,length(dat14$True)))
+    )
+str(newdat)
+
+library(ggplot2)
+library(viridis)
+vpal = viridis(4,end=0.9)
+
+p <- ggplot(newdat,aes(x=factor(which),y=response, fill=factor(which)))+geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))+
+    scale_fill_manual(values=vpal) +
+    scale_x_discrete(name="Response") +
+    scale_y_continuous(name="") +
+    theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
[…]

### 2019-01-24 revision/latex_template/author.tex
added revision files

--- 2018-11-14 manuscript/arxiv/nn-cytof.tex
@@
-\documentclass[conference]{IEEEtran}
+%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% sample root file for your "contribution" to a contributed volume
+%
+% Use this file as a template for your own input.
+%
+%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%========================================================================= PACKAGES
-\usepackage{amsmath,amssymb}
-\usepackage[mathscr]{euscript} %To be able to use script letters with \mathscr{}
-\usepackage{amsthm} %To be able to write Definitions, Theorems, etc.
-\usepackage{tikz} %To draw graphs.
-\usepackage{graphicx} % To scale equations
-\newcommand*{\Scale}[2][4]{\scalebox{#1}{\ensuremath{#2}}} % To scale equations
-\usepackage{framed} % To be able to frame theorems and stuff
-%\usepackage{float} %To be able to place figures exactly where I want with [H]
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{subfigure}
-\usepackage{longtable}
-\usepackage{MnSymbol}
-\usepackage{xcolor}  \definecolor{shadecolor}{rgb}{.95,.95,.95}  %To put a shaded region
-\usepackage[numbers,square, comma, sort&compress]{natbib}	 %Make the references [1-8], rather than [1]-[8], as with \usepackage{cite}
-\renewcommand*{\bibfont}{\footnotesize} %Size of the references; only needed with natbib
-%\usepackage{makeidx} %To create an index
-%\makeindex
-\usepackage[hidelinks]{hyperref} %To be able to use links inside the document; [hidelinks removes the ugly red boxes]
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[graybox]{svmult}
 
-%==================================================== ALGORITHM SETUP
-\usepackage[ruled,vlined]{algorithm2e}
+% choose options for [] as required from the list
+% in the Reference Guide
 
-%========================================================================= THEOREMS, LEMMAS, DEFINITIONS, ETC.
-\newtheorem{myDefinition}{Definition}
-\newtheorem{myTheorem}{Theorem}
-\newtheorem{myLemma}{Lemma}
-\newtheorem{myCorollary}{Corollary}
-\newtheorem{myProposition}{Proposition}
-\newtheorem{myExample}{Example}
-\newtheorem{myRemark}{Remark}
-\newtheorem{myConjecture}{Conjecture}
-\newtheorem{myQuestion}{Question}
-\newtheorem{myProblem}{Problem}
+\usepackage{mathptmx}       % selects Times Roman as basic font
+\usepackage{helvet}         % selects Helvetica as sans-serif font
+\usepackage{courier}        % selects Courier as typewriter font
+\usepackage{type1cm}        % activate if the above 3 fonts are
+                            % not available on your system
+%
+\usepackage{makeidx}         % allows index generation
+\usepackage{graphicx}        % standard LaTeX graphics tool
+                             % when including figure files
+\usepackage{multicol}        % used for the two-column index
+\usepackage[bottom]{footmisc}% places footnotes at page bottom
 
-%========================================================================= STUFF TO DRAW GRAPHS
-\tikzstyle{every edge}=  [draw]
-\tikzstyle{vertex} = [draw,circle,minimum size=1pt]
-\tikzstyle{label} = [minimum size=.1pt,font=\scriptsize]
-\tikzstyle{title} = [minimum size=.25cm,font=\small]
+% see the list of further useful packages
+% in the Reference Guide
 
-%========================================================================= SYMBOLS
-%\newcommand{\var}[1]{\mathsf{#1}}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\wt}[1]{\widetilde{#1}}
-\newcommand{\careful}{\textcolor{red}}
-\newcommand{\comment}{\textcolor{blue}}
+\makeindex             % used for the subject index
+                       % please use the style svind.ist with
+                       % your makeindex program
 
-%Miscelaneous
-%\def \R{\mathbb{R}}
-%\def \P{\mathsf{P}}
-%\def \T{\mathsf{T}}
-%\def \c{\mathsf{c}}
-\def \spn{{\rm span}}
-\def \rnk{{\rm rank}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%Scalars
-\def \r{{\hyperref[rDef]{r}}}
-
-%text
-\def \as{{\hyperref[asDef]{{\rm a.s.}}}}
-%Assumptions
-\def \Cone{{\hyperref[ConeDef]{{\rm {\bf C1}}}}}
-%Theorems
-\def \mainThm{{\hyperref[mainThm]{Theorem \ref{mainThm}}}}
-%Sections
-\def \modelSec{{\hyperref[modelSec]{Section \ref{modelSec}}}}
-%Figures
-\def \distributionFig{{\hyperref[distributionFig]{Figure \ref{distributionFig}}}}
-%Algorithms
-\def \validationAlg{{\hyperref[validationAlg]{Algorithm \ref{validationAlg}}}}
-
-\newcounter{@inst}
-\newcounter{@auth}
-\newcounter{auco}
-\newdimen\instindent
-\newbox\authrun
-\newtoks\authorrunning
-\newtoks\tocauthor
-\newbox\titrun
-\newtoks\titlerunning
-\newtoks\toctitle
-
-\def\clearheadinfo{\gdef\@author{No Author Given}%
-                   \gdef\@title{No Title Given}%
-                   \gdef\@subtitle{}%
-                   \gdef\@institute{}%
-                   \gdef\@thanks{}%
-                   \global\titlerunning={}\global\authorrunning={}%
-                   \global\toctitle={}\global\tocauthor={}}
-
-\def\institute#1{\gdef\@institute{#1}}
-
-\def\institutename
- {\begingroup
- \if!\@institute!\else
- \def\thanks##1{\unskip{}}%
- \def\iand{\\[5pt]\let\and=\nand}%
- \def\nand{\ifhmode\unskip\nobreak\fi\ $\cdot$ }%
- \let\and=\nand
- \def\at{\\\let\and=\iand}%
- \footnotetext[0]{\kern-\bibindent
- \ignorespaces\@institute}\vspace{5dd}\fi
- \endgroup
- }%
-
-% correct bad hyphenation here
-%\hyphenation{op-tical net-works semi-conduc-tor}
+%% track changes
+\usepackage[normalem]{ulem}
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}
+\newcommand{\missing}[1]{\textcolor{red}{\textbf{#1}}}
 
@@
 
-\title{Prediction of functional markers of mass cytometry data via
+\title*{Prediction of functional markers of mass cytometry data via
   deep learning}
@@
 
-\begin{abstract}
-Recently, there has been an increasing interest in the analysis of
-flow cytometry data, which involves measurements of a set of surface
-and functional markers across hundreds and thousands of cells. These
-measurements can often be used to differentiate various cell types and
-there has been a rapid development of analytic approaches for
-achieving this. However, in spite of the fact that measurements are
-available on such a large number of cells, there have been very
-limited advances in deep learning approaches for the analysis of flow
-cytometry data. Some preliminary work has focused on using deep
-learning techniques to classify cell types based on the cell protein
-measurements. In a first of its' kind study, we propose a novel deep
-learning architecture for predicting functional markers in the cells
-given data on surface markers. Such an approach is expected to
-automate the measurement of functional markers across cell samples,
-provided data on the surface markers are available, that has important
-practical advantages. We validate and compare our approach with
-competing machine learning methods using a real flow cytometry
-dataset, and showcase the improved prediction performance of the deep
-learning architecture.
-\end{abstract}
+\abstract{Recently, there has been an increasing interest in the analysis of flow cytometry data, which involves measurements of a set of surface and functional markers across hundreds and thousands of cells.
+These measurements can often be used to differentiate various cell types and there has been a rapid development of analytic approaches for achieving this.
+However, in spite of the fact that measurements are available on such a large number of cells, there have been very limited advances in deep learning approaches for the analysis of flow cytometry data.
+Some preliminary work has focused on using deep learning techniques to classify cell types based on the cell protein measurements.
+In a first of its' kind study, we propose a novel deep learning architecture for predicting functional markers in the cells given data on surface markers.
+Such an approach is expected to automate the measurement of functional markers across cell samples, provided data on the surface markers are available, that has important practical advantages.
+We validate and compare our approach with competing machine learning methods using a real flow cytometry dataset, and showcase the improved prediction performance of the deep learning architecture.}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[scale=0.35]{figures/surface-hist.pdf}
-\includegraphics[scale=0.35]{figures/functional-hist.pdf}
-\includegraphics[scale=0.35]{figures/scatterplot.pdf}
-\caption{Exploratory plots of surface and functional markers. The
-  histograms show a biased pattern, and the scatterplot shows non
-  linearity, both violations of crucial assumptions in standard
-  regression models.}
-\label{exploratory}
-\end{figure*}
-
 \subsection{Pre-processing}
@@
 \ref{exploratory}).
+
+\begin{figure}
+\includegraphics[scale=0.3]{figures/surface-hist.pdf}
+\includegraphics[scale=0.3]{figures/functional-hist.pdf}
+\includegraphics[scale=0.3]{figures/scatterplot.pdf}
+\caption{Exploratory plots of surface and functional markers. The
+  histograms show a biased pattern, and the scatterplot shows non
+  linearity, both violations of crucial assumptions in standard
+  regression models.}
+\label{exploratory}
+\end{figure}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[scale=0.4]{figures/neuralNet.eps}
-\caption{Neural network for predicting functional markers (18
-  responses) from surface markers (15 predictors) with 4 hidden
-  layers with 90, 90, 45 and 45 nodes each.}
-\label{nn}
-\end{figure*}
-
 % https://github.com/pluskid/Mocha.jl/blob/master/docs/user-guide/solver.rst
@@
 package \texttt{Mocha}\cite{Bezanson2017,mocha}.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.4]{figures/neuralNet.eps}
+\caption{Neural network for predicting functional markers (18
+  responses) from surface markers (15 predictors) with 4 hidden
+  layers with 90, 90, 45 and 45 nodes each.}
+\label{nn}
+\end{figure}
 
@@
 \centering
-\includegraphics[scale=0.45]{figures/vector-mse.pdf}
-\includegraphics[scale=0.45]{figures/time.pdf}
+\includegraphics[scale=0.35]{figures/vector-mse.pdf}
+\includegraphics[scale=0.35]{figures/time.pdf}
 \caption{Left: Vector MSE (equation \ref{vectorMSEformula}) for all
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[scale=0.35]{figures/violin.pdf} \hspace{0.25cm}
-\includegraphics[scale=0.35]{figures/response1.pdf}
+\includegraphics[scale=0.3]{figures/violin.pdf} \hspace{0.25cm}
+\includegraphics[scale=0.3]{figures/response1.pdf}
 \hspace{0.25cm}
-\includegraphics[scale=0.35]{figures/response14.pdf}
+\includegraphics[scale=0.3]{figures/response14.pdf}
 \caption{Left: Violin plot for four functional markers (responses).
@@
 \label{violin}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}
 \centering
@@
 
-% %======================================================= MAIN THEOREM
-% \begin{framed}
-% \begin{myTheorem}
-% \label{mainThm}
-% Write Theorem
-% \end{myTheorem}
-% \end{framed}
-
 \end{document}

### 2019-02-07 revision/answers2reviewers.md
starting revision: new results for plots, answers to reviwers

# Answers to reviewers

We thank the reviewers for such insightful comments which had certainly improved the manuscript greatly.

## From reviewer 1

1. "For the deep learning framework, there is not too much novelty. In my opinion, the paper is a pioneer of practitioner of the deep learning in their specific research area."
yes, we are not inventing neural networks, we are using them in this area where they have not been used before. we added this in the intro/discussion

2. "When they split the data into training, validation, test datasets, it’s better to follow the cross-validation schema, that is, try different repetition and report the average results to reduce the randomness."
definitely! we changed the analysis:  first the data was split into training set (1000,000 samples) and testing set (223,228 samples), and the training set was used to perform the 10-fold cross validation. Based on the average MSE from the cross validation, we chose the model with the smallest average MSE. Then we used all the training samples to obtain the final model and reported the prediction MSE of the testing samples.

3. "For the performance, we can see that the margin between NN and Random forest or linear model is not too much. SO, a better deep learning framework is still desired."
answer: in deep learning margins in performance can be relative. As we are comparing accuracy in prediction, random forest and NN are both quite competitive, with the former losing in computation time, rather than prediction accuracy. Other input is welcome. we added this in the intro/discussion

## From reviewer 2

1. "The study did not compare with support vector regression (SVR), citing “computational time constraints”. It is somewhat surprising since SVR is not a time-consuming algorithm comparing to deep learning.  Although the paper doesn’t have to do new work, some discussions along these lines could be added" 
It is our understanding that even when SVM are widely used in a variety of scenarios, SVR have not been as adopted in big data settings given its complexity.
One possible reason for the slow computation speed for SVR is if we are fitting a non-linear SVR then it could involve a radial basis function that involves computing a nXn kernel matrix where n is the number of samples which is very large for our application.

2. "The quality of the figures can be improved. Some labels in the figure are in light color, which is not easy to see. Some graphs (e.g. two in Fig. 2) have no axis labels."
Done, and figure 2 cannot have axis as it is the neural network architecture

3. "It would be good if the data used in this study can be made available to the public so that it can be served as a benchmark for studies in the community."
From Peng: The data is already publically available from the original paper that published the experiment, although one needs to learn a little bit about Cytobank to be able to get the data. Here is the link: http://reports.cytobank.org/1/v1

### 2019-02-07 script/plots.r

@@ -82,10 +82,18 @@ dev.off()
 
 ## Results copied from analysis-comb.jl, but
 ## later we will save them to files
+## For revision: results copied from analysis-comb-cv.jl
 
-vecloss = c(6.7148, 11.5280, 5.9503, 5.6447)
+## Original:
+vecloss = c(6.7148, 11.5280, 5.9503, 5.6447) ## search for vec_loss
+time = c(1.98,63.36,826.83,181.82) ##search for toc
+
+## Revision:
+vecloss = c(6.7147, 11.4857, 5.9399, 5.6474) ##search for vec_loss
+time = c(1.905,99.95,1093.24,195.64) ##search for toc
+
 method = c("Linear Model","Decision Tree", "Random Forest", "Neural Network")
-time = c(1.98,63.36,826.83,181.82)
+
 
 
 df = data.frame(vecloss=vecloss, method=method)
@@ -141,10 +149,17 @@ dev.off()
 ##########################################################################3
 method = c("Linear Model","Decision Tree", "Random Forest", "Neural Network")
 
+## Original: search for pt_loss
 lm = c(0.2308, 0.4301, 0.5954, 0.2353, 0.2490, 0.2427, 0.2414, 0.3983, 0.2016, 0.3520, 0.2792, 0.4322, 0.3795, 0.9793, 0.4539, 0.4587, 0.2931, 0.2624)
 dt = c(0.4207, 0.7510, 0.8740, 0.4556, 0.4644, 0.4460, 0.4363, 0.7468, 0.3831, 0.5550, 0.4696, 0.6693, 0.6602, 1.5796, 0.7696, 0.8004, 0.5239, 0.5227)
 rf = c(0.2183, 0.3881, 0.4480, 0.2374, 0.2429, 0.2321, 0.2278, 0.3863, 0.2005, 0.2869, 0.2441, 0.3433, 0.3425, 0.8015, 0.3982, 0.4095, 0.2725, 0.2703)
 nn = c(0.206722,0.358628,0.426626,0.225359,0.233064,0.221218,0.217177,0.366514,0.190379,0.273522,0.231513,0.327084,0.325995,0.756497,0.379003,0.389693,0.25933,0.256416)
+
+## Revision: search for pt_loss
[…]

### 2019-02-07 revision/answers2reviewers.md
modified manuscript according to reviews; only missing a few plots

@@
 1. "For the deep learning framework, there is not too much novelty. In my opinion, the paper is a pioneer of practitioner of the deep learning in their specific research area."
-yes, we are not inventing neural networks, we are using them in this area where they have not been used before. we added this in the intro/discussion
+We agree!
+Neural networks are widely studied models that have been applied to a variety of real-life datasets.
+We are not proposing new model developments for this area of neural network, we are applying existing neural network tools to the case of flow cytometry area which has not fully adopted the deep learning schemes yet.
+We added these points in the introduction/discussion
 
@@
 "When they split the data into training, validation, test datasets, it’s better to follow the cross-validation schema, that is, try different repetition and report the average results to reduce the randomness."
-definitely!
-we changed the analysis:  first the data was split into training set (1000,000 samples) and testing set (223,228 samples), and the training set was used to perform the 10-fold cross validation.
+This is a great point!
+We have changed the analysis accordingly: first the data was split into training set (1000,000 samples) and testing set (223,228 samples), and the training set was used to perform the 10-fold cross validation.
 Based on the average MSE from the cross validation, we chose the model with the smallest average MSE.
@@
 3. "For the performance, we can see that the margin between NN and Random forest or linear model is not too much. SO, a better deep learning framework is still desired."
-answer: in deep learning margins in performance can be relative.
+We agree!
+Sometimes in deep learning comparisons, margins in performance can be relative.
 As we are comparing accuracy in prediction, random forest and NN are both quite competitive, with the former losing in computation time, rather than prediction accuracy.
-Other input is welcome.
-we added this in the intro/discussion
+We added this discussion to the manuscript.
 
@@
 Although the paper doesn’t have to do new work, some discussions along these lines could be added"
-It is our understanding that even when SVM are widely used in a variety of scenarios, SVR have not been as adopted in big data settings given its complexity.
-One possible reason for the slow computation speed for SVR is if we are fitting a non-linear SVR then it could involve a radial basis function that involves computing a nXn kernel matrix where n is the number of samples which is very large for our application.
+It is our understanding that even when SVM are widely used in a variety of scenarios, SVR models have not been successfully adopted in big data settings given their complexity.
+One possible reason for the slow computation speed for SVR is that non-linear SVR involve a radial basis function that computes a n x n kernel matrix where n is the number of samples which is very large for our application.
+We have added these details and some references to the paper.
 
 2. "The quality of the figures can be improved. Some labels in the figure are in light color, which is not easy to see. Some graphs (e.g. two in Fig. 2) have no axis labels."
-Done, and figure 2 cannot have axis as it is the neural network architecture
+We have increased the font and changed the color of the figure labels. Figure 2 cannot have axis as it represents the neural network architecture.
 
 3. "It would be good if the data used in this study can be made available to the public so that it can be served as a benchmark for studies in the community."
-From Peng: The data is already publically available from the original paper that published the experiment, although one needs to learn a little bit about Cytobank to be able to get the data.
-Here is the link: http://reports.cytobank.org/1/v1
+The data is already publically available from the original paper that published the experiment. Here is the link: http://reports.cytobank.org/1/v1
+We added a section about Data Sharing in the manuscript.

### 2019-02-07 revision/latex_template/author.tex
modified manuscript according to reviews; only missing a few plots

@@
 
-\abstract{Recently, there has been an increasing interest in the analysis of flow cytometry data, which involves measurements of a set of surface and functional markers across hundreds and thousands of cells.
-These measurements can often be used to differentiate various cell types and there has been a rapid development of analytic approaches for achieving this.
-However, in spite of the fact that measurements are available on such a large number of cells, there have been very limited advances in deep learning approaches for the analysis of flow cytometry data.
-Some preliminary work has focused on using deep learning techniques to classify cell types based on the cell protein measurements.
-In a first of its' kind study, we propose a novel deep learning architecture for predicting functional markers in the cells given data on surface markers.
-Such an approach is expected to automate the measurement of functional markers across cell samples, provided data on the surface markers are available, that has important practical advantages.
-We validate and compare our approach with competing machine learning methods using a real flow cytometry dataset, and showcase the improved prediction performance of the deep learning architecture.}
+\abstract{Recently, there has been an increasing interest in the
+  analysis of flow cytometry data, which involves measurements of a
+  set of surface and functional markers across hundreds and thousands
+  of cells. These measurements can often be used to differentiate
+  various cell types and there has been a rapid development of
+  analytic approaches for achieving this. However, in spite of the
+  fact that measurements are available on such a large number of
+  cells, there have been very limited advances in deep learning
+  approaches for the analysis of flow cytometry data. Some preliminary
+  work has focused on using deep learning techniques to classify cell
+  types based on the cell protein measurements. In a first of its'
+  kind study, we propose a novel deep learning architecture for
+  predicting functional markers in the cells given data on surface
+  markers. Such an approach is expected to automate the measurement of
+  functional markers across cell samples, provided data on the surface
+  markers are available, that has important practical advantages. We
+  validate and compare our approach with competing machine learning
+  methods using a real flow cytometry dataset, and showcase the
+  improved prediction performance of the deep learning architecture.}
 
@@
 interactions, among many others, by becoming a major discovery force
-in science due to the unprecedented accuracy in prediction.  Moreover,
-deep learning approaches have shown accurate performance on genomics
-and biomedical
-applications\cite{Ciresan2013,Denas2013,Fakoor2013,Leung2014,CruzRoa2014,Li2017,Mobadersany2018}.
-
-Furthermore, CyTOF data is perfectly suited for deep learning methods.
-On one side, identify markers define a cell type (e.g., B cell, T
-cell, monocytes, MSC), and on the other side, expressions of
-functional markers identify the cell's activity (e.g., quiescent,
-secreting cytokines, proliferating, apoptosis). Since CyTOF technology
-allows for the simultaneous measurement of a large number of protein
-markers, most CyTOF studies measure both identity markers and
-functional markers, providing data for supervised learning tools, like
-neural networks.  In addition, each CyTOF run typically collects data
-on $10^6$ cells, creating an ideal large dataset in which the number
-of samples (cells) is orders of magnitude larger than the number of
+in science due to the unprecedented accuracy in prediction.
+\claudia{Moreover}{However}, \claudia{}{despite} deep learning
+approaches \claudia{have shown}{showing} accurate performance on
+genomics and biomedical
+applications\cite{Ciresan2013,Denas2013,Fakoor2013,Leung2014,CruzRoa2014,Li2017,Mobadersany2018},
+\claudia{}{the CyTOF community has not fully adopted these
+  approaches.}  \claudia{Furthermore}{In fact, it turns out that}
+CyTOF data is perfectly suited for deep learning methods.  On one
+side, identify markers define a cell type (e.g., B cell, T cell,
+monocytes, MSC), and on the other side, expressions of functional
+markers identify the cell's activity (e.g., quiescent, secreting
+cytokines, proliferating, apoptosis). Since CyTOF technology allows
+for the simultaneous measurement of a large number of protein markers,
+most CyTOF studies measure both identity markers and functional
+markers, providing data for supervised learning tools, like neural
+networks.  In addition, each CyTOF run typically collects data on
+$10^6$ cells, creating an ideal large dataset in which the number of
+samples (cells) is orders of magnitude larger than the number of
 variables (markers). Deep learning methods are particularly suited for
@@
 
-In terms of motivation, there are two main reasons to predict the functional markers from surface markers in CyTOF data: 1) monetary and time cost, and 2) technical limit of the total number of markers CyTOF can measure, which is currently around 50 protein markers.
-That is, if we can accurately predict some functional markers based on surface markers, there is no longer the need to include those functional markers in the staining panel (experimental design), and thereby freeing up channels to measure more surface markers or additional functional markers that cannot be predicted.
+In terms of motivation, there are two main reasons to predict the
+functional markers from surface markers in CyTOF data: 1) monetary and
+time cost, and 2) technical limit of the total number of markers CyTOF
+can measure, which is currently around 50 protein markers. That is, if
+we can accurately predict some functional markers based on surface
+markers, there is no longer the need to include those functional
+markers in the staining panel (experimental design), and thereby
+freeing up channels to measure more surface markers or additional
+functional markers that cannot be predicted.
 
-Here, we explore neural network models to predict functional markers
-(internal phosphoproteins) with identify markers (cell surface
-proteins), and compare its performance in terms of accuracy and speed
-to other standard statistical approaches like regression, and random
-forests.  We show that neural networks improve prediction of
-functional markers, making them a powerful alternative to the usual
-regression techniques.
+Here, we \claudia{}{present the first work to} explore neural network
+models to predict functional markers (internal phosphoproteins) with
+identify markers (cell surface proteins). We compare the performance
+of neural networks in terms of accuracy and speed to other standard
+statistical approaches like regression, and random forests.  We show
+that neural networks improve prediction of functional markers, making
+them a powerful alternative to the usual regression techniques,
+\claudia{}{thus, providing quantitive evidence that deep learning
+  methods can enrich the existing research landscape of CyTOF big
+  data.}
 
@@
 2) decision trees, and 3) random trees. Due to computational time
-constraints, we could not fit a support vector regression (SVR)
-model. We compared the performance of the four approaches by computing
-the mean square error (MSE) of the predicted responses.
+constraints, we could not fit a support vector regression (SVR) model
+(more in Discussion). We compared the performance of the four
+approaches by computing the mean square error (MSE) of the predicted
+responses.
 
-To fit the models, we divided the data into training set, validation
-set and testing set. The training set was used to fit each of the four
-models. The validation set was used to determine the best setup
-(tuning parameters) of each model in terms of MSE. Finally, the test
-set was used to compare the MSE across methods.
+To fit the models, we divided the data into training set,
+\claudia{validation set}{} and testing set. The training set was used
+to \claudia{fit each of the four models}{perform a 10-fold cross
+  validation scheme to choose the best model fit for each method. We
+  chose the model fit with the smallest average MSE}. \claudia{The
+  validation set was used to determine the best setup (tuning
+  parameters) of each model in terms of MSE. Finally, the test set was
+  used to compare the MSE across methods.}{Then, we used all the
+  training samples to estimate the final model for each method
+  (e.g. model and tuning parameters), and reported the prediction MSE
+  of each method based on the testing samples.}
 
@@
 markers (responses) and 15 surface markers (predictors), which we
-divided as follows: 750,000 rows as training set, 250,000 rows as
-validation set, and 223,228 rows as testing set.
+divided as follows: \claudia{750,000}{ 1,000,000} rows as training set,
+\claudia{250,000 rows as validation set}{}, and 223,228 rows as
+testing set.
 
@@
 architectures, activation functions, regularization coefficients,
-solver methods, momentum policies, and learning rates with 50,000 maximum epochs. The best
-network has four layers (see figure \ref{nn}) with 90, 90, 45 and 45
-nodes. The network uses hyperbolic tangent as activation function,
-regularization coefficient of 0.0001, momentum policy fixed at 0.8, inverse-decay learning
-rate policy with base learning rate, gamma and power parameters at 0.01, 0.0001, 0.75. We used Adam solver\cite{Kingma2014}, an
-algorithm for first-order gradient-based optimization of stochastic
-objective functions, based on adaptive estimates of lower-order
-moments, instead of Stochastic Gradient Descent (SGD) as the former
-showed increased accuracy. All networks were trained using the julia
-package \texttt{Mocha}\cite{Bezanson2017,mocha}.
+solver methods, momentum policies, and learning rates with 50,000
+maximum epochs. The best network has four layers (see figure \ref{nn})
+with 90, 90, 45 and 45 nodes. The network uses hyperbolic tangent as
+activation function, regularization coefficient of 0.0001, momentum
+policy fixed at 0.8, inverse-decay learning rate policy with base
+learning rate, gamma and power parameters at 0.01, 0.0001, 0.75. We
+used Adam solver\cite{Kingma2014}, an algorithm for first-order
+gradient-based optimization of stochastic objective functions, based
+on adaptive estimates of lower-order moments, instead of Stochastic
+Gradient Descent (SGD) as the former showed increased accuracy. All
+networks were trained using the julia package
+\texttt{Mocha}\cite{Bezanson2017,mocha}.
 
@@
 
-In this work, we showed that a neural network model outperforms
-standard statistical approaches like linear regression and random
-forest in the prediction of functional markers from surface markers
-for CyTOF data. Neural networks were also faster and more efficient
-than random forests, which make them a more viable choice for big
-datasets.
+\claudia{}{To the best of our knowledge, our work presents the first
+  neural network model for the prediction of functionality in CyTOF
+  big data.}  In this work, we showed that a neural network model
+outperforms standard statistical approaches like linear regression and
+random forest in the prediction of functional markers from surface
+markers for CyTOF data. Neural networks were also faster and more
+efficient than random forests, which make them a more viable choice
+for big datasets.
 
@@
 among responses by fitting a network for the whole response vector.
+\claudia{}{It is important to note that the while the neural network
+  model shows better performance than the other methods, this edge is
+  still quite modest. Random forest models also represent a suitable
+  choice for prediction under non-linear assumptions. However, we show
+  here that when scalability also matters, the neural network model
+  outperforms random forests (in computation time), while providing
+  more accurate predictions. We claim that neural networks are better
+  suited for big CyTOF data than random forests, but more work is
+  needed to corroborate this claims.}
 
@@
 
-The accuracy in the prediction of functional markers from surface markers has economic and computational advantages, for example, considering the limitation to the total number of markers CyTOF can measure, which is currently around 50 protein markers.
-Being able to predict functional markers from surface markers could allow for different types of staining panels which could measure more surface markers, or focus on functional markers not so easily predicted.
+The accuracy in the prediction of functional markers from surface
+markers has economic and computational advantages, for example,
+considering the limitation to the total number of markers CyTOF can
+measure, which is currently around 50 protein markers. Being able to
+predict functional markers from surface markers could allow for
+different types of staining panels which could measure more surface
+markers, or focus on functional markers not so easily predicted.
+
+\claudia{}{Initially, we were surprised by the lack of scalability of
+  SVR models. However, given the non-linearity and optimization
+  complexity\cite{Smola2003, Welling2004}, it appeared that our
+  dataset was too large for the computation of the kernel matrix,
+  which is $n \times n$ where $n$ is the number of samples (more than
+  1 million in our case). More work is needed to assess the
+  performance of SVR under smaller datasets to assess the accuracy of
+  such models and the limits in scalability.}
 
@@
 
+\vspace{0.5cm}
+\noindent \textbf{Acknowledgements:} We thank the editor as well as
+two anonymous reviewers whose suggestions greatly improved the
+manuscript.
+
+\vspace{0.5cm}
+\noindent \textbf{Data Sharing:} The data is publicly available
+through the original publications\cite{Bendall2011,Qiu2011} in the
+following website: \url{http://reports.cytobank.org/1/v1}.
+
 % For future work, So, we have N experiments: N+1 matrices 100k by 50: B

### 2019-02-07 script/plots.r

@@ -106,11 +106,11 @@ p <- ggplot(df,aes(x=reorder(method,-vecloss),y=vecloss,fill=method))+geom_bar(s
             plot.title = element_text(hjust=0.5, size=rel(1.8)),
             axis.title.x = element_text(size=rel(1.8)),
             axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-            axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-            axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+            axis.text.x = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
+            axis.text.y = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
             ##legend.text=element_text(size=rel(1.2)), legend.title=element_text(size=rel(1.5)),
             panel.background = element_blank(),
-            axis.line = element_line(colour = "grey")
+            axis.line = element_line(colour = "black")
             ) +
     xlab("") + ylab("Vector MSE") + guides(fill=FALSE) +coord_flip()
 
@@ -132,11 +132,11 @@ p <- ggplot(df,aes(x=reorder(method,-time),y=time,fill=method))+geom_bar(stat="i
             plot.title = element_text(hjust=0.5, size=rel(1.8)),
             axis.title.x = element_text(size=rel(1.8)),
             axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-            axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-            axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+            axis.text.x = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
+            axis.text.y = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
             ##legend.text=element_text(size=rel(1.2)), legend.title=element_text(size=rel(1.5)),
             panel.background = element_blank(),
-            axis.line = element_line(colour = "grey")
[…]

### 2019-02-07 revision/answers2reviewers.md
finished resubmission

@@
 Neural networks are widely studied models that have been applied to a variety of real-life datasets.
-We are not proposing new model developments for this area of neural network, we are applying existing neural network tools to the case of flow cytometry area which has not fully adopted the deep learning schemes yet.
+We are not proposing new model developments for the area of neural network; we are applying existing neural network tools to the case of flow cytometry area which has not fully adopted the deep learning schemes yet.
 We added these points in the introduction/discussion

### 2019-02-07 revision/latex_template/author.tex
finished resubmission

@@
 \claudia{}{the CyTOF community has not fully adopted these
-  approaches.}  \claudia{Furthermore}{In fact, it turns out that}
+  methods yet.}  \claudia{Furthermore}{In fact, it turns out that}
 CyTOF data is perfectly suited for deep learning methods.  On one

### 2019-02-07 script/plots.r

@@ -214,11 +214,18 @@ dev.off()
 ## Plots per response
 ## We will focus in response 1 (good), 3, 8, 14 (bad)
 
-## aqui voy: Waiting for the revised files after CV model fit
+## Original
 dat1 = read.table("../results/Predictions/Test_1.txt", header=TRUE)
 dat3 = read.table("../results/Predictions/Test_3.txt", header=TRUE)
 dat8 = read.table("../results/Predictions/Test_8.txt", header=TRUE)
 dat14 = read.table("../results/Predictions/Test_14.txt", header=TRUE)
+
+## Revision
+dat1 = read.table("../results/Test_prediction/Test_1_revision.txt", header=TRUE)
+dat3 = read.table("../results/Test_prediction/Test_3_revision.txt", header=TRUE)
+dat8 = read.table("../results/Test_prediction/Test_8_revision.txt", header=TRUE)
+dat14 = read.table("../results/Test_prediction/Test_14_revision.txt", header=TRUE)
+
 
 newdat = data.frame(response = c(dat1$True,dat3$True,dat8$True,dat14$True),
     which = c(rep(1,length(dat1$True)),rep(3,length(dat3$True)),rep(8,length(dat8$True)),rep(14,length(dat14$True)))
@@ -237,15 +244,16 @@ p <- ggplot(newdat,aes(x=factor(which),y=response, fill=factor(which)))+geom_vio
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
         axis.title.x = element_text(size=rel(1.8)),
         axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=90, hjust=.5, vjust=.5, face="plain"),
+        axis.text.x = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="black", size=rel(1.7), angle=0, hjust=.5, vjust=.5, face="plain"),
         legend.text=element_text(size=rel(1.2)), legend.title=element_blank(),
         panel.background = element_blank(),
-        axis.line = element_line(colour = "grey"),
+        axis.line = element_line(colour = "black"),
         legend.position = "none"
[…]

### 2019-02-08 revision/answers2reviewers-sk.md
added changes suprateek

--- 2019-02-07 revision/answers2reviewers.md
@@
 We are not proposing new model developments for the area of neural network; we are applying existing neural network tools to the case of flow cytometry area which has not fully adopted the deep learning schemes yet.
+To our knowledge, there has been a limited development of neural network approaches for prediction of cell proteins based on flow cytometry data, and the proposed method fills this gap.
+We hope that this work will encourage the development of future deep learning approaches in flow cytometry.
 We added these points in the introduction/discussion
@@
 
-3. "For the performance, we can see that the margin between NN and Random forest or linear model is not too much. SO, a better deep learning framework is still desired."
+SK comment: we need to insert what changes in the results we observed, if any, when we used the cross-validation approach versus the results reported earlier.
+
+3. "For the performance, we can see that the margin between NN and Random forest or linear model is not too much. So, a better deep learning framework is still desired."
 We agree!
@@
 As we are comparing accuracy in prediction, random forest and NN are both quite competitive, with the former losing in computation time, rather than prediction accuracy.
+One key advantage of the NN over random forest is that the former is able to handle a larger number of covariates, and hence is more scalable to higher dimensional feature spaces compared to the random forest.
+We expect the computation time under random forest to slow down even further as the number of covariates increases.
 We added this discussion to the manuscript.
+
+SK comment: did the performance of the NN improve relative to random forest under the cross-validation criteria?
 
@@
 1.
-"The study did not compare with support vector regression (SVR), citing “computational time constraints”.
+"The study did not compare with support vector regression (SVR), citing computational time constraints.
 It is somewhat surprising since SVR is not a time-consuming algorithm comparing to deep learning.
-Although the paper doesn’t have to do new work, some discussions along these lines could be added"
+Although the paper doesn't have to do new work, some discussions along these lines could be added"
 It is our understanding that even when SVM are widely used in a variety of scenarios, SVR models have not been successfully adopted in big data settings given their complexity.
-One possible reason for the slow computation speed for SVR is that non-linear SVR involve a radial basis function that computes a n x n kernel matrix where n is the number of samples which is very large for our application.
+One possible reason for the slow computation speed for SVR is that non-linear SVR involve a radial basis function that computes a n x n kernel matrix at each iteration, where n is the number of samples (cells) which is very large for our application.
+Computing and storing such a matrix is likely to slow down computation and this is likely what we experienced when we tried to implement the SVR approach.
 We have added these details and some references to the paper.
@@
 2. "The quality of the figures can be improved. Some labels in the figure are in light color, which is not easy to see. Some graphs (e.g. two in Fig. 2) have no axis labels."
-We have increased the font and changed the color of the figure labels. Figure 2 cannot have axis as it represents the neural network architecture.
+Thank you for this comment. We have increased the font and changed the color of the figure labels. Figure 2 cannot have axis as it represents the neural network architecture.
 
 3. "It would be good if the data used in this study can be made available to the public so that it can be served as a benchmark for studies in the community."
-The data is already publically available from the original paper that published the experiment. Here is the link: http://reports.cytobank.org/1/v1
+Yes we fully agree. The data is already publically available from the original paper that published the experiment. Here is the link: http://reports.cytobank.org/1/v1
 We added a section about Data Sharing in the manuscript.

### 2019-03-11 revision/latex_template/author-track.tex
updated revised manuscript

--- 2019-02-07 revision/latex_template/author.tex
@@
   work has focused on using deep learning techniques to classify cell
-  types based on the cell protein measurements. In a first of its'
+  types based on the cell protein measurements. In a first of its
   kind study, we propose a novel deep learning architecture for

### 2019-03-13 script/analysis-comb-cv.jl
Update analysis-comb-cv.jl

@@ -20,7 +20,7 @@ functional=["141-pPLCgamma2","150-pSTAT5", "152-Ki67","154-pSHP2","151-pERK1/2",
             "171-pBtk/Itk","172-pS6","174-pSrcFK","176-pCREB","175-pCrkL"]
 dfpred = vcat(df1[Symbol.(surface)],df2[Symbol.(surface)],df3[Symbol.(surface)],df4[Symbol.(surface)],df5[Symbol.(surface)])
 dfresp = vcat(df1[Symbol.(functional)],df2[Symbol.(functional)],df3[Symbol.(functional)],df4[Symbol.(functional)],df5[Symbol.(functional)])
-## randomize 1,223,228 rows into three sets: train (1,000,000 rows), test (223,228 rows)
+## randomize 1,223,228 rows into two sets: train (1,000,000 rows), test (223,228 rows)
 srand(5312)
 rand_seq = randperm(1223228)
 dfpredTrain = convert(Array{Float64,2}, dfpred[rand_seq[1:1000000],:])
